Childhood exposure to non-persistent endocrine disruptors, glucocorticosteroids, and neurodevelopment: A study based on the parametric g-formula

### 2024-01-04 output/paper/manuscript.qmd
Rendered manuscript to send to Martine.

We are not aware of prior studies specifically investigating the effects of exposure to \acr{edcs} in relation to \acr{hitrtse}. The literature on \acr{edcs} and neurodevelopment in children has mostly focused on \acr{oops} [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @CartierWarembourgLeManer-Idrissi:2016; YuDuChiou:2016], phthalate metabolites [@HuangChenSu:2015; @HuangTsaiChen:2017; @KimHongShin:2017; @VilmandBeckBilenberg:2023], and \acr{bpa} [@TewarAuingerBraun:2016; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019], in relation to \acr{adhd} [@BouchardBellingerWright:2010; @TewarAuingerBraun:2016; @YuDuChiou:2016; @LiZhangKuang:2018], and intelligence scales [@Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @HuangChenSu:2015; @CartierWarembourgLeManer-Idrissi:2016; @HuangTsaiChen:2017; @KimHongShin:2017; @VilmandBeckBilenberg:2023]. Few studies have looked into different classes of \acr{edcs} [@ShoaffCoullWeuve:2020 in relation with the Conners Attention Deficit Scale and the Behavior Assessment System for Children; @OhKimKannan:2023 in relation with \acr{adhd} symptoms]. Overall, and consistent with our results, these studies seem to provide further evidence of the negative effects of several \acr{edcs} on neurodevelopment in children. While not all these studies have investigated effect modification by sex, it seems that these negative effects are stronger in males. A major limitation of these studies is the reliance on spot urine samples, that might not be representative of long-term exposures.

>>>>

We are not aware of prior studies specifically investigating the effects of exposure to \acr{edcs} in relation to \acr{hitrtse}. The literature on \acr{edcs} and neurodevelopment in children has mostly focused on \acr{opps} [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @CartierWarembourgLeManer-Idrissi:2016; YuDuChiou:2016], phthalate metabolites [@HuangChenSu:2015; @HuangTsaiChen:2017; @KimHongShin:2017; @VilmandBeckBilenberg:2023], and \acr{bpa} [@TewarAuingerBraun:2016; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019], in relation to \acr{adhd} [@BouchardBellingerWright:2010; @TewarAuingerBraun:2016; @YuDuChiou:2016; @LiZhangKuang:2018], and intelligence scales [@Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @HuangChenSu:2015; @CartierWarembourgLeManer-Idrissi:2016; @HuangTsaiChen:2017; @KimHongShin:2017; @VilmandBeckBilenberg:2023]. Few studies have looked into different classes of \acr{edcs} [@ShoaffCoullWeuve:2020 in relation with the Conners Attention Deficit Scale and the Behavior Assessment System for Children; @OhKimKannan:2023 in relation with \acr{adhd} symptoms]. Overall, and consistent with our results, these studies seem to provide further evidence of the negative effects of several \acr{edcs} on neurodevelopment in children. While not all these studies have investigated effect modification by sex, it seems that these negative effects are stronger in males. A major limitation of these studies is the reliance on spot urine samples, that might not be representative of long-term exposures.
